---
title: "Project proposal"
author: "Jack, Luke, Dexter"
output: html_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
```

## 1. Introduction

The data has been collected from Major League Baseball's Statcast program. It contains pitcher data from 2021 to 2024, detailing the outcome of each pitch thrown. Each observation is the summary statistic of each pitcher in a specific year, where it demonstrates the outcomes of all pitches thrown. Each variable represents the outcome of a pitch, or event such as a pickoff attempt. Our research question is: "Is there a correlation between pitcher actions(Pitches, Pickoff Attempts, etc.) and pitcher success measured in wins and losses."

## 2. Data

```{r data review}

getwd()
baseball <- read_csv("../data/pitching_data_2021-2024.csv")
glimpse(baseball)

```

## 3. Data analysis plan

Our plan is to understand what pitch combinations make a pitcher better or worse. This can be done by looking at pitch combinations and batter success against a pitcher. Visualizing this will present an image of success. We will also look at time trends to see how pitchers change their pitch combinations as they age in order to stay successful. The visualizations below demonstrate that there is differentiation among the data in terms of strikeouts and that there is a probable correlation between innings pitched and Earned Run Average-the average number of runs allowed over 9 innings-(ERA), where as more innings are pitched, ERA decreases. Our visualizations can look at sucess at different ages and look to see how throwing different combinations of pitches can be successful. We can also look at how success differs between wins and ERA. Therefore we can understand which pitchers may be winning because they have a powerful offense on their team, wehreas some may be failing eventhough they have a good ERA because their offense does not generate enough runs.

```{r preliminary visualizations}
library(ggplot2)
#Summary Statistics
summary(baseball)
#Histogram of strikeouts
ggplot(baseball, aes(x = strikeout)) +
  geom_histogram(binwidth = 10, fill = "blue", color = "black") +
  labs(title = "Distribution of Strikeouts", x = "Strikeouts", y = "Frequency") 

#ERA vs Innings pitched
ggplot(baseball, aes(x = p_formatted_ip, y = p_era)) +
  geom_point(color = "red", alpha = 0.6) +
  labs(title = "Innings Pitched vs ERA", x = "Innings Pitched", y = "ERA")
```



## 4. Data Ethics Review

The data ethics review section will be introduced in a separate class and is not part of the original proposal deadline.

