---
title: "Project proposal"
author: "Jack, Luke, Dexter"
output: html_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
```

## 1. Introduction

The data has been collected from Major League Baseball's Statcast program. It contains pitcher data from 2021 to 2024, detailing the outcome of each pitch thrown. Each observation is the summary statistic of each pitcher in a specific year, where it demonstrates the outcomes of all pitches thrown. Each variable represents the outcome of a pitch, or event such as a pickoff attempt. Our research question is: "Is there a correlation between pitcher actions and pitcher success measured in wins and losses."

## 2. Data

```{r}

getwd()
baseball <- read_csv("../data/pitching_data_2021-2024.csv")
glimpse(baseball)

```

## 3. Data analysis plan

Our plan is to understand what pitch combinations make a pitcher better or worse. This can be done by looking at pitch combinations and batter success against a pitcher. Visualizing this will present an image of success. We will also look at time trends to see how pitchers change their pitch combinations as they age in order to stay successful.



## 4. Data Ethics Review

The data ethics review section will be introduced in a separate class and is not part of the original proposal deadline.

